---
title: "SDS315 - Homework 02"
author: "Rachel Chavez - rcc3342"
date: "2025-01-23"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

# Homework 02

## **Question 1**

```{r read1}

library(tidyverse)
library(ggplot2)
library(kableExtra)

profs <- read_csv("profs.csv")

```

```{r A1}

ggplot(profs) + geom_histogram(aes(x = eval), bins = 30, fill = 'lavender', col = 'black') + labs(title = "UT Austin Professors' Average Course Evaluation Scores", x = "Score", y = "Count") + theme_bw()

```

Pictured above is the distribution of counts of professors' average
scores (x-axis), rated by students. The distribution of average course
evaluation's scores has a left skew with a peak at about 4, implying
that a majority of professors have mean scores at about 3.5 with the
median being at 4.

```{r B1}

ggplot(profs) + geom_boxplot(aes(x = eval, y = native), fill = 'lavender') + labs(title = "Average Course Evaluation Scores by Native Language", x = "Score", y = "Is the Professor a Native English Speaker?") + scale_y_discrete(labels = c("yes" = "Yes", "no" = "No")) + xlim(0, 5) + theme_bw()

```

The difference of average scores (x-axis) between professors who are
native English speakers as opposed to not is displayed above. The
average score for native English speakers is higher than that of
non-native speakers, but it also exhibits a wider range. The 25th
percentile for native speakers is lower than that of non-native
speakers, while the 75th percentile is higher, but it appears generally
that non-native English speakers are rated lower.

```{r C1}

ggplot(profs) + geom_histogram(aes(x = eval), bins = 40, fill = 'lavender', col = "black") +
  # facet into 2 rows
  facet_wrap(~gender, nrow = 2) + 
  labs(title = "Average Course Evaluation Scores by Gender", x = "Score", y = "Count") +
  theme_bw()

```

Pictured above are the average score distributions of female (top) and
male (bottom) professors. The distribution among male professors has a
strong left skew while the female distribution is more symmetrical
(though still left skewed), implying that males generally receive less
lower scores.

```{r D1}

ggplot(profs) + geom_point(aes(y = eval, x = beauty), col = "purple4") +
  labs(title = "Average Course Scores by Average Beauty Rating", y = "Score", x = "Beauty Rating") +
  theme_bw()

```

The scatterplot above examines the relationship between how attractive
students consider the professor (x-axis) compared to their average
scores (y-axis). Less attractive professors had a wider range of scores
while more attractive professors had a narrower scope of higher scores
with the correlation between the two being weak but positive at
`r cor(profs$eval, profs$beauty)`.

## **Question 2**

```{r read2}

bikeshare <- read_csv("bikeshare.csv")

```

```{r A2}

avg_rentals <- bikeshare %>%
  group_by(hr) %>%
  summarise(avg_boardings = mean(total))

ggplot(avg_rentals) + geom_line(aes(x = hr, y = avg_boardings), linewidth =0.5) +
  labs(title = "Average Bike Rentals by Hour", x = "Hour (24:00)", y = "Average Rentals") + 
  theme_bw()

```

This line graph shows the average number of bike rentals in Washington
DC as the day progresses. From midnight to about 5:00AM, there are very
few rentals, on average, but soon after, they begin steadily increasing
to above 200 until they peak at about 3:30PM and remain high until night
rolls around at 8:00PM, and rentals decrease steadily back down.

```{r B2}

avg_rentals <- bikeshare %>%
  group_by(hr, workingday) %>%
  summarise(avg_boardings = mean(total), .groups = 'drop')

ggplot(avg_rentals) + geom_line(aes(y = avg_boardings, x = hr), linewidth =0.5) +
  # changing facets with new labels
  facet_wrap(~workingday,
             labeller = labeller(workingday = c("0" = "Weekend", "1" = "Weekday"))) +
  labs(title = "Average Bike Rentals vs. Hour by Weekend", x = "Hour (24:00)", y = "Average Rental Count") +
  theme_bw()

```

This figure depicts average hourly bike rentals (with the hour being on
the x-axis and the rental counts being on the y-axis), comparing weekend
data to weekday data. On weekdays, bike rentals significantly spike at
around 8:00AM and again at 5:00PM, likely related to work schedules,
while weekends have a unimodal distribution, staying above 200 rentals
per hour from about 10:00AM til 8:00PM.

```{r C2}

# Plot C: a faceted bar plot showing average ridership (y) during the 9 AM hour by weather situationcode (weathersit, x), faceted according to whether it is a working day or not. (Remember that you can focus on a specific subset of rows of a data set using filter.)

ninedata <- bikeshare %>% 
  filter(hr == 9) %>%
  group_by(weathersit, workingday) %>%
  summarise(avg_ride = mean(total), .groups = 'drop')

ggplot(ninedata) + geom_col(aes(x = weathersit, y = avg_ride), fill = 'lightblue', col = 'black') +
  facet_wrap(~workingday, labeller = labeller(workingday = c("0" = "Weekend", "1" = "Weekday"))) +
  labs(title = "Average Ridership at 9AM by Weather Severity", x = "Weather Severity", y = "Ridership Count") +
  theme_bw()

```

The graph depicts the average ridership of bikes (y-axis) on both
weekend and weekdays by the weather conditions (x-axis). Working days
have significantly more bike rentals than non-working days, and both
have strong right skews, seeing most bike rentals in clear skies
(*weather situation 1*) with very little levels during intense weather
(*weather situation 3*).

## **Question 3**

```{r read3}

capmetro_UT <- read_csv("capmetro_UT.csv")

capmetro_UT = mutate(capmetro_UT,
  day_of_week = factor(day_of_week,
  levels=c("Mon", "Tue", "Wed","Thu", "Fri", "Sat", "Sun")),
  month = factor(month,
  levels=c("Sep", "Oct","Nov")))

```

```{r A3}

capmetro_UTn = capmetro_UT %>%
  select(hour_of_day, day_of_week, boarding, month) %>%
  group_by(hour_of_day, month, day_of_week) %>%
  mutate(avg_boarding = mean(boarding)) %>%
  ungroup()

ggplot(capmetro_UTn) + geom_line(aes(x=hour_of_day, y=avg_boarding, col = month)) +
  facet_wrap(~day_of_week) + 
  labs(title = "Seasonal Average CapMetro UT Boardings by Hour of Day", x = "Hour of the Day (24:00)", y = "Number of Boardings") +
  theme_bw()

```

The graphs above show the distribution of average boardings (y-axis) by
the hour (y-axis) throughout each day of the week with each line
representing a different month (being September, October, and November).

The peak hour of business appears to stay broadly consistent across each
weekday, but the weekends have a much less obvious peak that appears to
occur near the end of the day. All peaks are roughly around 5:00PM to
8:00PM. The number of boardings on Mondays in September may be affected
by Labor Day falling on Monday and lowering the amount of
transportation. Similarly, during November, when Thanksgiving occurs,
many people travel out of town or stay inside with family, leading them
to rely less on public transportation.

```{r B3}

ggplot(capmetro_UT) + geom_point(aes(x = temperature, y = boarding, col = weekend)) +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.5), text=element_text(size=7)) +
  labs(title = "CapMetro Boardings by Temperature for 6:00-21:00", x = "Temperature(F)", y = "Boarding Counts") +
  #facet_grid(~hour_of_day)
  facet_wrap(~hour_of_day)

```

The above figure depicts the boarding of UT students on CapMetro with
each image representing an hour of the day, the x-axis being the
increasing temperature, and the y-axis being the number of boarding
students.

With hour of day and weekend status held constant, temperature does not
seem to have a noticeable effect on the number of UT students riding the
bus. As the temperature raises, there does not appear to be a
significant increase or decrease in number of boardings as the slope
remains somewhat flat across all instances.

## **Question 4**

```{r read4}

billboard = read_csv("billboard.csv")

```

```{r A4}

top_ten = billboard %>%
  select(c(performer, song, weeks_on_chart)) %>%
  
  group_by(performer, song) %>%
  summarise(total_weeks = n(), .groups = 'drop') %>%
  
  arrange(desc(total_weeks)) %>%
  slice(1:10)

my.tab = kbl(top_ten, caption = "Top 10 Most Popular Songs by Weeks on Billboard",
             col.names = c("Artist", "Song Title", "Total Weeks"))
kable_styling(my.tab, bootstrap_options = "striped")
# ADD A CAPTION TO THE TABLE YOU BUFFOON
```

The table above depicts the top 10 most popular songs from 1958 to 2021,
measured by the total number of weeks each song was on the Billboard Top
100. It has three columns: artist name, song name, and weeks on
Billboard.

```{r B4}

uni_year = billboard %>%
  filter(year != 1958 & year !=2021) %>%
  group_by(year) %>%
  summarize(unique_songs = length(unique(song_id)), .groups = 'drop')

ggplot(uni_year) + geom_line(aes(y = unique_songs, x = year)) +
  labs(title = "Musical Diversity on Billboard Top 100, 1959-2020", x = "Year", y = "Number of Unique Songs") + 
  theme_bw()

```

This graph shows the number of unique songs (y-axis) that appeared in
the Billboard Top 100 by each year (x-axis). There was a large drop
beginning at the end of 1900's, but the number of unique songs steadily
rose again in the 2000's before raising to about 800 songs in 2020.

```{r C4}

top_19 = billboard %>%
  select(c(song, performer, weeks_on_chart)) %>%
  # only pick songs that were on the charts for 10 or more weeks
  filter(weeks_on_chart >= 10) %>%
  # group by performer AND song in case multiple songs have the same title
  group_by(performer, song) %>%
  # count how many performer/song pairs there are
  summarise(weeks10 = n(), .groups = 'drop') %>%
  # combine the artists' total 10 week songs to see their total number of hits
  group_by(performer) %>% 
  summarise(top_songs = n(), .groups = 'drop') %>%  
  # throw out all artists who have 30 or less hits
  filter(top_songs >= 30)

ggplot(top_19) +  geom_col(aes(x = performer, y = top_songs), fill = 'lightblue', col = 'black') +
  theme_bw() +
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.5), text=element_text(size=7)) +
  labs(title = "Artists with Most Popular Songs", x = "Artist", y = "Number of 10-week Hit Songs")
```

19 artists since 1958 have had at least 30 songs that were on the
Billboard Top 10 for at least ten weeks. The graph depicts the artists
on the x-axis with the number of 10-week hit songs on the y-axis. The
artist with the most number of hits was Elton John with 52 songs that
were hits for ten weeks.
