---
title: "Homework 08"
author: "Rachel Chavez - rcc3342 - https://github.com/ray-chav/SDS-315/blob/main/HW08.Rmd"
date: "2025-04-09"
output:
  pdf_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

# Homework 08

```{r library}

library(tidyverse)
library(kableExtra)
library(ggplot2)
library(minpack.lm)
library(mosaic)

```

# Problem 1: regression warm up

Below, a linear model is fitted to data of people's creatinine levels (mL/minute) and their ages.

```{r read1}

creatinine <- read_csv('creatinine.csv')

```

## A) What creatinine clearance rate should we expect for a 55-year-old? Explain briefly (one or two sentences + equations) how you determined this.

By using the lm() function in R, we can build a linear function that minimizes each data point's distance to its line. When we want to find a person's expected creatinine level based off of their age, we can do so by plugging in our independent age parameter into the model.

```{r A1}

# build linear model
creat_model <- lm(creatclear ~ age, data = creatinine)

# now specifically check for the 55 year old
#print(predict(creat_model, newdata = data.frame(age = 55)))
# 113.723

```

Based off of this model, for a 55-year-old, we can estimate their
creatinine clearance rate at
`r round(predict(creat_model, newdata = data.frame(age = 55)),3)`
mL/minute.

## B) How does creatinine clearance rate change with age? (This should be a single number whose units are ml/minute per year.) Explain briefly (one or two sentences) how you determined this.

The change in creatinine clearance rate with age is represented by the
slope of the regression model, which is approximately -0.619 mL/min per
year.

```{r B1}

#summary(creat_model)
# The change in creatinine clearance rate with age is represented by the slope of the regression model, which is -0.61982 mL/min per year.

```

## C) Whose creatinine clearance rate is healthier (higher) for their age: a 40-year-old with a rate of 135, or a 60-year-old with a rate of 112? Explain briefly (a few sentences + equations) how you determined this

To determine which creatinine clearance rate is healthier for their age,
we compare their actual rate to the predicted value from a regression model, based on their age. To determine where somebody's clearance rate is in comparison to their expected value, we can do the following equation: $Difference=ObservedClearance-PredictedClearance$, where the observed clearance is the given creatinine clearance rate, and the predicted clearance is the output of our linear function when given the observed age.

For the 40-year-old, their creatinine clearance rate of 135 is 11.98
mL/minute higher than the predicted value for someone of that age.

For the 60-year-old, their creatinine clearance rate of 112 is 1.38
mL/minute higher than the predicted value for someone of that age.

Thus, the 40-year-old has a higher relative creatinine clearance rate
compared to their age, indicating that their rate is healthier
for their age.

```{r C1}

# 40 year old
#print(135 - predict(creat_model, newdata = data.frame(age = 40)))
# 11.98

# 60 year old
#print(112 - predict(creat_model, newdata = data.frame(age = 60)))
# 1.38

```

# Problem 2: Modeling disease growth

The below is an analysis of COVID-19 cases in Italy and Spain based off of their daily recorded deaths and the days elapsed since the country's first COVID-19 related death.

## 1) An estimated growth rate and doubling time for Italy, with 95% bootstrapped confidence intervals for each.

Below is a table of summary statistics for this data set. From it, we see
that the average deaths for Italy and Spain were 318.718 and 292.464 deaths, respectively,
occuring 19.0 and 13.5 days after the first recorded death, on average.

By fitting an exponential growth model to Italy and Spain's COVID-19
data, we estimate our observed growth rate to be 0.183 and 0.276,
respectively. Thus, the observed doubling times for this sample were 3.8
and 2.5.

```{r A2}

# reading in data
covid_data <- read_csv("covid.csv")

# filter data for Italy and Spain
italy_data <- filter(covid_data, country == "Italy")
spain_data <- filter(covid_data, country == "Spain")

# tibble of summary statistics
tab <- covid_data %>%
  group_by(Country = country) %>%
  summarize(Average_Deaths = round(mean(deaths), 3), Average_Days = mean(days_since_first_death))

kable_styling(kable(
  tab,
  caption = "Summary Statistics of COVID-19 Deaths"
))

# fit exponential growth models for Italy and Spain
# deaths = exp(intercept + growth_rate * days_since_first_death)
italym <- lm(log(deaths) ~ days_since_first_death, data = italy_data)
spainm <- lm(log(deaths) ~ days_since_first_death, data = spain_data)

# grab model summaries
# coef(italym) %>% round(0)
# coef(spainm) %>% round(0)

# find b (growth rates) for both models
# rounded to nearest third
italy_growth_rate <- round(coef(italym)[2], 3)
spain_growth_rate <- round(coef(spainm)[2], 3)

# doubling time = ln(2) / growth_rate
# rounded to nearest tenth
italy_doubling_time <- round(log(2) / italy_growth_rate, 1)
spain_doubling_time <- round(log(2) / spain_growth_rate, 1)

# print both growth rates and doubling time out
print(paste("Italy Growth Rate:", italy_growth_rate, "Doubling Time:", italy_doubling_time))
print(paste("Spain Growth Rate:", spain_growth_rate, "Doubling Time:", spain_doubling_time))

```

Now, by bootstrapping 10,000 samples from our original sample of Italy's
COVID cases, we can construct a confidence interval for both of these
statistics.

```{r}

#--------
# BOOT STRAPPING!!

# bootstrap a bunch of samples from italy
# and then build a lm for each one
italy_bootstrap <- do(10000) * {
  model <- lm(log(deaths) ~ days_since_first_death, data = mosaic::resample(italy_data))
  coef(model)[2]  # extract the slope (growth rate)
}

# confidence interval
italy_ci <- confint(italy_bootstrap, value = 0.95)

# print it out os i can look at it teehee
print(paste("Italy Growth Rate 95% CI: ", 
    round(italy_ci$upper, 3), " to ", round(italy_ci$lower, 3)))

# calculate doubling time for all of our bootstrapped growth rates
italy_doubling_time <- log(2) / italy_bootstrap

# confidence interval
italy_dt_ci <- confint(italy_doubling_time, value = 0.95)

# print out doubling time estimates
print(paste("Italy Doubling Time 95% CI: ", 
    round(italy_dt_ci$lower, 1), " to ", round(italy_dt_ci$upper, 1)))

```

Using an exponential growth model fitted to Spain’s bootstrapped
samples, we estimate Italy's COVID-19 growth rate to fall between 0.201 and 0.165 increase in
deaths per day, with 95% confidence. Additionally, its doubling time for
COVID deaths was between 3.7 to 3.8 days.

## 2) An estimated growth rate and doubling time for Spain, with 95% bootstrapped confidence intervals for each.

Based on our confidence interval built on bootstrapping, the estimated daily growth rate of deaths was between 0.233 and 0.322 per day, with 95% confidence. Based on this, the estimated
doubling time of daily deaths in Spain was between 2.4 and 2.5 days,
indicating that the death count was doubling roughly every two and a half days.

```{r B2}

# same deal as above but with Spain
spain_bootstrap <- do(10000) * {
  model <- lm(log(deaths) ~ days_since_first_death, data = mosaic::resample(spain_data))
  coef(model)[2]  # extract the slope (growth rate)
}

# confidence interval
spain_ci <- confint(spain_bootstrap, value = 0.95)
print(paste("Spain Growth Rate 95% CI: ", 
    round(spain_ci$lower, 3), " to ", round(spain_ci$upper, 3)))

# now recalculate doubling time into a df from our bootstrapped growth rates
spain_doubling_time <- log(2) / spain_bootstrap

# new confidence interval
spain_dt_ci <- confint(spain_doubling_time, value = 0.95)
print(paste("Spain Doubling Time 95% CI: ", 
    round(spain_dt_ci$lower, 1), " to ", round(spain_dt_ci$upper, 1)))

```

## 3) A line graph showing reported daily deaths over time (using days_since_first_death, rather than calendar date, as the relevant time variable) in each country. Your line graph should have two lines, one for each country, distinguished by their color.

The graph shows the number of reported COVID-19 deaths over the number of days since each
country's first reported death in Italy and Spain. Each point represents the number of
deaths on a given day, and the two countries are distinguished by color.
The positive slope of each line indicates an upward, exponential trend in deaths during the early
stages of the pandemic.

The shaded regions around each country's trend line represent 95%
confidence intervals, being the range of uncertainty for the true
trend in deaths over days.

```{r C2}

# plotting
ggplot(covid_data) + 
  geom_point(aes(x=days_since_first_death, y=deaths, color=country), alpha = 0.3) +
  geom_smooth(aes(x=days_since_first_death, y=deaths, color=country),  method='lm')

```

# Problem 3: price elasticity of demand

To estimate the price elasticity of demand for milk, we used a power-law
model (log(Q)=log(K)+Blog(P)). We fit a linear
regression model with log(sales) as the response and log(price) as the
predictor. The estimated price elasticity for our initial sample is -1.6186 , indicating
that a 1% increase in milk price leads to an estimated 1.62% decrease in quantity demanded.

From our original sample, we can bootstrap 10,000 resamples in order to continuously remodel our data and recalculate our elasticity. 

```{r C3}

milk <- read_csv('milk.csv')

# Q = KP^B
# logQ = Blog(P)
# build elasticity model
milk_model <- lm(log(sales) ~ log(price), data = milk)

# grab our sample B
coef(milk_model)[2] %>% round(4)
# B = -1.6186

# milk bootstrap
milk_bootstrap <- do(10000) * {
  model = lm(log(sales) ~ log(price), data = mosaic::resample(milk))
  coef(model)[2]
}

# confidence interval for milk price elasticity
milk_ci <- confint(milk_bootstrap$log.price., value = 0.95)
#milk_ci <- confint(milk_bootstrap, value = 0.95)


# print confidence interval
print(paste("Milk Elasticity 95% CI: ", 
    round(milk_ci$`2.5%`, 4), " to ", round(milk_ci$`97.5%`, 4)))

```

To quantify uncertainty, we used bootstrapping with 1,000 resamples. The 95% confidence interval for the price elasticity of demand is between -1.7706 and -1.4603. This suggests that as the price of milk increases by 1.0%, the quantity demanded decreases by approximately 1.46% to 1.77%. Since the absolute value of the elasticity is greater than 1, this indicates that milk demand is elastic. Customers are relatively responsive to changes in price.
